Formulating RDF Constraints and \\ Validating RDF Data Generically

### 2015-04-20 paper/paper 2/paper 2.tex
simple constraints

For \emph{property constraints}, \textbf{\emph{left}} and \textbf{\emph{right property lists}} are specified.
The assignment of properties to these lists happens relative to the \emph{constraining element} - which may be an operator (e.g., $\sqsubseteq$ in case of \emph{object property paths}).
\emph{Object Property Paths}\footnote{Corresponds to the requirement \emph{R-55-OBJECT-PROPERTY-PATHS}} (also called \emph{complex role inclusion axioms} in \emph{DL})
express that, if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the \emph{author} of the book \emph{A-Brief-History-Of-Time} (\ms{authorOf(Stephen-Hawking, A-Brief-History-Of-Time)}) whose \emph{genre} is \emph{Popular-Science} ({\ms{genre(A-Brief-History-Of-Time, Popular-Science)}),
the following \emph{object property path} infers that \emph{Stephen-Hawking} is an \emph{author} of the \emph{genre} \emph{Popular-Science} (\ms{authorOfGenre(Stephen-Hawking, Popular-Science)}):

\begin{DL}
authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre 
\end{DL}

Thus, when mapped to the \emph{RDF-CV} (see table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are on the left side of the \emph{constraining element} \emph{$\sqsubseteq$},
and the property \emph{authorOfGenre} is on the right side:

@@

There are \emph{simple constraints} which cannot be expressed in DL such as \emph{literal pattern matching}, \emph{literal value comparison}, \emph{literal ranges}, \emph{default values}, and \emph{language tag cardinality} \cite{BoschNolleAcarEckert2015}.
% maybe additional constraint default values:
% ear shape of races, e.g. vulcans / teeth color
% mathematical operations

{\em Literal pattern matching}\footnote{Corresponds to {\em R-44-PATTERN-MATCHING-ON-RDF-LITERALS}} restricts literals to match given patterns.
The universal restriction \ms{Book $\sqsubseteq$ $\forall$ identifier.ISBN}, e.g., ensures that \emph{books} can only have valid ISBN identifiers, 
i.e. strings that match a given regular expression. 
The restriction of the datatype \emph{ISBN}, however, cannot be expressed in DL, but OWL 2 DL can be used anyway to express the \emph{literal pattern matching} constraint:

\begin{ex}
ISBN a RDFS:Datatype ; owl:equivalentClass [ a RDFS:Datatype ;
    owl:onDatatype xsd:string ; owl:withRestrictions ([ xsd:pattern "^\d{9}[\d|X]\$" ])] .
\end{ex}
\tb{ToDO: delete the slash in front of the dollar sign from the output of the code fragment above (in the stated pattern)?}
The first OWL 2 axiom explicitly declares {\em ISBN} to be a datatype. %\er{Why start from second and then first? Also better we say "owl axiom" when we talk axiom of owl, to prevent ambiguity.}\tb{resolved}
The second OWL 2 axiom defines {\em ISBN} as an abbreviation for a datatype restriction on {\em xsd:string}. 
The datatype {\em ISBN} can be used just like any other datatype like in the universal restriction above.
%There are multiple use cases associated with the requirement to match literals according to given patterns (\emph{Literal Pattern Matching}\footnote{Corresponds to \emph{R-44-PATTERN-MATCHING-ON-RDF-LITERALS}}).
%The enterprise vessel, e.g.,  can only have the registry numbers "NCC-1701", "NCC-1701-A", "NCC-1701-B", "NCC-1701-C", "NCC-1701-D", or "NCC-1701-E".
%The universal restriction can be represented in DL:
%\ms{Enterprise $\sqsubseteq$ $\forall$ registryNumber.RegistryNumber}.
%The restriction of the datatype \ms{RegistryNumber}, however, cannot be expressed in DL, but OWL 2 DL can be used anyway to express the literal pattern matching constraint:
%
%\begin{ex}
%RegistryNumber
    %a rdfs:Datatype ; owl:equivalentClass [ a rdfs:Datatype ;
        %owl:onDatatype xsd:string ;
        %owl:withRestrictions ( [ xsd:pattern "NCC-1701([-][A-E])?" ] ) ] .
%\end{ex}
%The second axiom defines \ms{RegistryNumber} as an abbreviation for a datatype restriction on \ms{xsd:string}. 
%The first axiom explicitly declares \ms{RegistryNumber} to be a datatype. 
%The datatype \ms{RegistryNumber} can be used just like any other datatype like in the universal restriction above.
The \emph{literal pattern matching} constraint validates \emph{ISBN} literals according to the stated regular expression causing constraint violations for triples not matching the regular expression. 
%\ms{Janeway commandsEnterprise Voyager} and \ms{Voyager registryNumber "NCC-74656"\textasciicircum{}\textasciicircum{}RegistryNumber}, 
%but not for the triples \ms{Picard commandsEnterprise Enterprise} and \ms{Enterprise registryNumber "NCC-1701-E"\textasciicircum{}\textasciicircum{}RegistryNumber}.
In table \ref{tab:simple-constraint-not-expressible-in-dl)}, the \emph{simple constraint} (\emph{literal pattern matching}) and the \emph{complex constraint} (\emph{universal restriction}) are mapped to the \emph{RDF-CV}.

@@

\begin{table}
  \scriptsize
  \sffamily
  \vspace{0cm}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      property & Book & identifier & - & ISBN & $\forall$ & - \\
      class & ISBN & - & - & xsd:string & regex & '\string^\text{$\backslash$d$\{9\}$[$\backslash$d$\mid$X]}\$' \\
		\end{tabular}
	\caption{Simple Constraint (Not Expressible in DL)}
	\label{tab:simple-constraint-not-expressible-in-dl)}
\end{table} %\tb{Erman, can you please format the regular expression in the table above correctly?}
The \emph{literal pattern matching} constraint introduces the new \emph{constraining element} \emph{regex}.
For this new \emph{constraining element}, the validation must be implemented.
This has to be done once for each \emph{generic constraint type} which cannot be expressed in DL.

>>>>

For \emph{property constraints}, \textbf{\emph{left}} and \textbf{\emph{right property lists}} are specified.
The assignment of properties to these lists happens relative to the \emph{constraining element} - which may be an operator (e.g., $\sqsubseteq$ in case of \emph{object property paths}).
\emph{Object Property Paths}\footnote{Corresponds to the requirement \emph{R-55-OBJECT-PROPERTY-PATHS}} (also called \emph{complex role inclusion axioms} in \emph{DL})
express that, if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the \emph{author} of the book \emph{A-Brief-History-Of-Time} (\ms{authorOf(Stephen-Hawking, A-Brief-History-Of-Time)}) whose \emph{genre} is \emph{Popular-Science} ({\ms{genre(A-Brief-History-Of-Time, Popular-Science)}),
the following \emph{object property path} infers that \emph{Stephen-Hawking} is an \emph{author} of the \emph{genre} \emph{Popular-Science} (\ms{authorOfGenre(Stephen-Hawking, Popular-Science)}):
\begin{DL}
authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre 
\end{DL}
Thus, when mapped to the \emph{RDF-CV} (see table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are on the left side of the \emph{constraining element} \emph{$\sqsubseteq$},
and the property \emph{authorOfGenre} is on the right side:

@@

There are \emph{simple constraints} which cannot be expressed in \emph{DL} such as \emph{literal pattern matching}, \emph{literal value comparison}, \emph{literal ranges}, \emph{default values}, and \emph{language tag cardinality} \cite{BoschNolleAcarEckert2015}.
% maybe additional constraint default values:
% ear shape of races, e.g. vulcans / teeth color
% mathematical operations

{\em Literal pattern matching}\footnote{Corresponds to {\em R-44-PATTERN-MATCHING-ON-RDF-LITERALS}} restricts literals to match given patterns.
The following \emph{universal restriction}, e.g., ensures that \emph{books} can only have valid \emph{ISBN} identifiers, 
i.e., strings that match a given regular expression:
\begin{DL}
\ms{Book $\sqsubseteq$ $\forall$ identifier.ISBN} 
\end{DL}
Even though, the restriction of the datatype \emph{ISBN} cannot be expressed in \emph{DL}, \emph{OWL 2 DL} can be used to express the \emph{literal pattern matching} constraint:

\begin{ex}
ISBN a RDFS:Datatype ; owl:equivalentClass [ a RDFS:Datatype ;
    owl:onDatatype xsd:string ; 
    owl:withRestrictions ([ xsd:pattern "^\d{9}[\d|X]$" ])] .
\end{ex}\$$
The first OWL 2 axiom explicitly declares {\em ISBN} to be a datatype. %\er{Why start from second and then first? Also better we say "owl axiom" when we talk axiom of owl, to prevent ambiguity.}\tb{resolved}
The second OWL 2 axiom defines {\em ISBN} as an abbreviation for a datatype restriction on {\em xsd:string}. 
The datatype {\em ISBN} can be used just like any other datatype like in the \emph{universal restriction} above.
%There are multiple use cases associated with the requirement to match literals according to given patterns (\emph{Literal Pattern Matching}\footnote{Corresponds to \emph{R-44-PATTERN-MATCHING-ON-RDF-LITERALS}}).
%The enterprise vessel, e.g.,  can only have the registry numbers "NCC-1701", "NCC-1701-A", "NCC-1701-B", "NCC-1701-C", "NCC-1701-D", or "NCC-1701-E".
%The universal restriction can be represented in DL:
%\ms{Enterprise $\sqsubseteq$ $\forall$ registryNumber.RegistryNumber}.
%The restriction of the datatype \ms{RegistryNumber}, however, cannot be expressed in DL, but OWL 2 DL can be used anyway to express the literal pattern matching constraint:
%
%\begin{ex}
%RegistryNumber
    %a rdfs:Datatype ; owl:equivalentClass [ a rdfs:Datatype ;
        %owl:onDatatype xsd:string ;
        %owl:withRestrictions ( [ xsd:pattern "NCC-1701([-][A-E])?" ] ) ] .
%\end{ex}
%The second axiom defines \ms{RegistryNumber} as an abbreviation for a datatype restriction on \ms{xsd:string}. 
%The first axiom explicitly declares \ms{RegistryNumber} to be a datatype. 
%The datatype \ms{RegistryNumber} can be used just like any other datatype like in the universal restriction above.
The \emph{literal pattern matching} constraint validates \emph{ISBN} literals according to the regular expression causing constraint violations for triples which do not not match. 
%\ms{Janeway commandsEnterprise Voyager} and \ms{Voyager registryNumber "NCC-74656"\textasciicircum{}\textasciicircum{}RegistryNumber}, 
%but not for the triples \ms{Picard commandsEnterprise Enterprise} and \ms{Enterprise registryNumber "NCC-1701-E"\textasciicircum{}\textasciicircum{}RegistryNumber}.
In table \ref{tab:simple-constraint-not-expressible-in-dl)}, the \emph{simple constraint} (\emph{literal pattern matching}) and the \emph{complex constraint} (\emph{universal restriction}) are mapped to the \emph{RDF-CV}:

@@

\begin{table}
  \scriptsize
  \sffamily
  \vspace{0cm}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      property & Book & identifier & - & ISBN & $\forall$ & - \\
      class & ISBN & - & - & xsd:string & regex & '\string^\text{$\backslash$d$\{9\}$[$\backslash$d$\mid$X]}\$' \\
		\end{tabular}
	\caption{Simple Constraint (Not Expressible in DL)}
	\label{tab:simple-constraint-not-expressible-in-dl)}
\end{table} %\tb{Erman, can you please format the regular expression in the table above correctly?}
The \emph{literal pattern matching} constraint type introduces the new \emph{constraining element} \emph{regex},
for which a validation mechanism has to be implemented.
Validation has to be implemented once for each \emph{generic constraint type} which is not expressible in \emph{DL}.
